Guidelines for reliable and reproducible functional enrichment analysis

### 2021-08-31 analysis.Rmd
manuscript editing
minor edits
fix journal names and generate pdfs

## Filter for studies included in QC2

```{r,xy}

@@

@@

@@

@@

@@

@@

cor.test(xm$scores , xm$cites, method="spearman")
```

>>>>

write.table(x,file="TableS2.tsv",quote=FALSE,sep="\t",col.names = TRUE,row.names = FALSE)

@@

names(res) <- gsub("RNA array","Gene expression array",names(res))

@@

pdf("images/hist1.pdf",width=4,height=3)
par(mar=c(5,5,3,1))
barplot(table(scores), xlab="analysis score",ylab="frequency")
dev.off()

@@

pdf("images/score_sjr1.pdf",width=4,height=3)
par(mar=c(4,4,3,2))
plot(xm$SJR,xm$scores, col = rgb(red=0,blue=0,green=0,alpha=0.2), pch=19,bty="n",
  xlab="SJR",ylab="score",main="Analysis scores vs. journal metrics")
abline(mylm1,col="red")
grid()
dev.off()

@@

pdf("images/score_highest1.pdf",width=3,height=4.5)
par(mar=c(5,8,3,1))
barplot(tail(jres$scores,20),names.arg = tail(jres$Journal,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="mean score",
        main = "Highest scoring",xlim=c(-4,0))
grid()
dev.off()

pdf("images/score_lowest1.pdf",width=3,height=4.5)
par(mar=c(5,8,3,1))
barplot(head(jres$scores,20),names.arg = head(jres$Journal,20),horiz=TRUE,las=1,cex.names = 0.7, xlab="mean score",
        main = "Lowest scoring",xlim=c(-5,0))
grid()
dev.off()

@@

pdf("images/mean_sjr1.pdf",width=3.5,height=3)
par(mar=c(5,5,3,1))
plot(jjres$SJR,jjres$scores,  col = rgb(red=0,blue=0,green=0,alpha=0.2) , pch=19,bty="n",
  xlab="SJR",ylab="mean score",main="Mean analysis score vs. journal metrics")
abline(mylm2,col="red")
dev.off()

@@

pdf("images/cites1.pdf",width=3.5,height=3)
par(mar=c(5,5,3,1))
plot(log2(xm$cites),xm$scores,col = rgb(red=0,blue=0,green=0,alpha=0.2) , pch=19,bty="n",
  xlab="log2(citations)",ylab="score",main="Analysis scores vs. log2(citations)")
abline(mylm2,col="red")
dev.off()

### 2021-08-31 manuscript.Rmd
fix journal names and generate pdfs

There were articles from 100 journals in the sample, with *PLoS One*, *Scientific Reports*, and *PeerJ* being the biggest contributors (Figure S1A).
There were 18 different omics types, with gene expression array and RNA-seq being the most popular (Figure S1B).
There were 31 different species under study, but human was the most common with 157 analyses (Figure S1C).

@@

We scored a total of 1630 analyses, revealing only a small fraction of articles obtained a satisfactory score of zero or higher.
Surprisingly, the analysis scores we generated did not correlate with journal level bibliographic metric, nor accrued citations of individual articles, showing no association between methodological rigour and traditional research impact metrics in this sample.
This suggests that methodological and reporting problems are not limited to lower ranked journals, but are a more general problem.

@@

5. Always report FDR or q-values (p-values that have been corrected for multiple tests).
This will reduce the chance of false positives when performing many tests simultaneously.
Report what method was used for p-value adjustment.
Bonferonni, FDR and FWER are examples.

@@

We thank Dr Nicholas C. Wong (Monash University) and Dr Antony Kaspi (Walter and Eliza Hall Institute) for comments on the manuscript.
This research was supported by use of the Nectar Research Cloud, a collaborative Australian research platform supported by the NCRIS-funded Australian Research Data Commons (ARDC).

@@

>>>>

There were articles from 96 journals in the sample, with *PLoS One*, *Scientific Reports*, and *PeerJ* being the biggest contributors (Figure S1A).
There were 18 different omics types, with gene expression array and RNA-seq being the most popular (Figure S1B).
There were 31 different species under study, but human was the most common with 157 analyses (Figure S1C).

@@

We scored a total of 1630 analyses, revealing only a small fraction of articles obtained a satisfactory score of zero or higher.
The analysis scores we generated did not correlate with journal or article metrics.
This suggests that methodological and reporting problems are not limited to lower ranked journals, but are a more general problem.

@@

5. Always report FDR or q-values (p-values that have been corrected for multiple tests).
This will reduce the chance of false positives when performing many tests simultaneously (Hung et al, 2012).
Report what method was used for p-value adjustment.
Bonferonni, FDR and FWER are examples.

@@

We thank XX for comments on the manuscript.
This research was supported by use of the Nectar Research Cloud, a collaborative Australian research platform supported by the NCRIS-funded Australian Research Data Commons (ARDC).

@@

Hung, J.-H., Yang, T.-H., Hu, Z., Weng, Z., & DeLisi, C. (2012). Gene set enrichment analysis: performance evaluation and usage guidelines. Briefings in Bioinformatics, 13(3), 281–291.
